---
title: "California Vaccinations in Kindergarten Students"
output: 
 flexdashboard::flex_dashboard:
   orientation: rows
   vertical_layout: scroll
   source_code: embed
   theme: lumen
runtime: shiny
---

```{r context="setup", include=FALSE}
library(shiny)
library(ggplot2)
library(dplyr)
library(flexdashboard)
library(plotly)
```

```{r}
vaccine_df <- read.csv(file="StudentData_clean.csv", header=TRUE, sep=",")
```


Inputs {.sidebar}
=======================================================================

```{r context="render"}
numericInput(inputId = "n_samp", 
             label = "Sample Size:", 
             min = 1, max = nrow(vaccine_df),
             value = 100)
```

```{r context="render"}
vaccine_sample <- reactive({
  req(input$n_samp)
  sample_n(vaccine_df, input$n_samp)
})

```

```{r context="render"}
checkboxGroupInput(inputId = 'school_type',
                   label = "School Type:",
                   choices = c('Public' = 'PUBLIC',
                               'Private' = 'PRIVATE'),
                   selected = c('PUBLIC', 'PRIVATE'))
```

```{r}
school_type_subset <- reactive({
  vaccine_sample() %>%
    filter(Type %in% input$school_type)
})
```


Vaccination Definitions:

* MMR: measles, mumps, and rubella 
* DTP: diphtheria, tetanus, and pertussis
* Polio: polio
* PBE: personanl belief exemption
* PME: personal medical exemptions


```{r context="render"}
selectInput(inputId = 'vacc_type',
                   label = "Vaccination:",
                   choices = c('MMR' = 'MMR',
                               'DTP' = 'DTP',
                               'Polio' = 'Polio',
                               'Personal Belief Exemption' = 'PBE',
                               'Medical Exemption' = 'PME'),
                   selected = 'MMR')
```


Note: Herd Imumnity is the resistance to the spread of a contagious disease within a population that results if a sufficiently high proportion of individuals are immune to the disease, especially through vaccination.  The herd immunity rate for measles of 95% was used in determining the gauge for "sucess."


Vaccination-Specific Analysis
=======================================================================


Row {data-height=175}
-----------------------------------------------------------------------

### Average Number of Students per School

```{r}
renderValueBox({
  avg_students <- mean(school_type_subset()[, 'Students'])
  valueBox(round(avg_students),  icon='fa-pencil')
})
```


### Percentage Vaccinated based on Herd Immunity Requirements

```{r}
renderGauge({
  avg_vacc <- mean(school_type_subset()[,input$vacc_type], na.rm = TRUE)
  gauge(round(avg_vacc), min = 0, max = 100, symbol = '%', gaugeSectors(
    success = c(95, 100), warning = c(90, 94), danger = c(0, 89)))
    })
```


Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Percent Vaccinated by School Type (1)

```{r context="server"}
output$Density <- renderPlotly({
  ggplotly(
    ggplot(school_type_subset(), 
           aes_string(x=input$vacc_type, fill='Type')) + 
      geom_density(alpha=.15) +
      labs(x="% of Childen Vaccinated",
           y="",
           fill="School Type",
           title="Percent of Kindergarteners Vaccinated"), 
    tooltip = c('x', 'fill')
    )
})


```

```{r context="render"}
plotlyOutput('Density')
```


### Percent Vaccinated by School Type (2)

```{r context="server"}

output$Violin <- renderPlotly({
  req(input$vacc_type)
  ggplotly(
    ggplot(school_type_subset(), aes_string(x='Type',
                                            y=input$vacc_type,
                                            fill='Type')) +
      geom_violin() +
      guides(fill=FALSE) +
      labs(x="% of Childen Vaccinated",
           y="",
           fill="Type",
           title="Percent of Kindergarteners Vaccinated"), 
    tooltip = c('y', 'x')
    )
  })
```

```{r context="render"}
plotlyOutput('Violin')
```


Row
-----------------------------------------------------------------------

### Percent Vaccinated by County

```{r context="server"}
output$County <- renderPlotly({
  ggplotly(
    ggplot(school_type_subset(), aes_string(x='County', 
                                            y=input$vacc_type, 
                                            fill='County')) + 
      geom_boxplot() + 
      guides(fill=FALSE) +
      theme(axis.text.x = element_text(angle=65))
  )
})
```

```{r context="render"}
plotlyOutput('County')
```


Overall Vaccination Data
=======================================================================

Row {data-height=175}
-----------------------------------------------------------------------

### Average Number of Students with MMR Vaccination

```{r}
renderValueBox({
  avg_MMR <- mean(school_type_subset()$MMR)
  valueBox(scales::percent(avg_MMR/100), icon='fa-syringe', color='orange')
})
```


### Average Number of Students with DTP Vaccination

```{r}
renderValueBox({
  avg_DTP <- mean(school_type_subset()$DTP)
  valueBox(scales::percent(avg_DTP/100), icon='fa-syringe', color='red')
})
```


### Average Number of Students with Polio Vaccination

```{r}
renderValueBox({
  avg_Polio <- mean(school_type_subset()$Polio)
  valueBox(scales::percent(avg_Polio/100), icon='fa-syringe', color='fuchsia')
})
```


### Average Number of Students with Personal Belief Exemption (PBE)

```{r}
renderValueBox({
  avg_PBE <- mean(school_type_subset()$PBE)
  valueBox(scales::percent(avg_PBE/100), icon='fa-syringe', color='green')
})
```


### Average Number of Students with Medical Exemption (PME)

```{r}
renderValueBox({
  avg_PME <- mean(school_type_subset()$PME)
  valueBox(scales::percent(avg_PME/100), icon='fa-syringe', color='blue')
})
```


Row {.tabset .tabset-fade data-height=700}
-----------------------------------------------------------------------

### Percent of Children Vaccinated by Year

```{r context="server"}
output$Bubbles <- renderPlotly({
  ggplotly(
    ggplot(school_type_subset(), aes(x=Year, size=Students)) +
      geom_jitter(aes(y=MMR, col='MMR'), width=.5, alpha=.5) +
      geom_jitter(aes(y=DTP, col='DTP'), width=.5, alpha=.5) +
      geom_jitter(aes(y=Polio, col='Polio'), width=.5, alpha=.5) +
      geom_jitter(aes(y=PBE, col='PBE'), width=.5, alpha=.5) +
      geom_jitter(aes(y=PME, col='PME'), width=.5, alpha=.5) +
      labs(x="Year",
           y="% of Childen Vaccinated",
           fill="School Type",
           title="Kindergarteners Vaccinated by Year"),
    tooltip = c("y", "x", 'size') 
    )
  })
```

```{r context="render"}
plotlyOutput('Bubbles')
```


### Kindergarten Vacination Data Table

```{r context="server"}
output$DataTable <- DT::renderDataTable({
  DT::datatable(school_type_subset(), 
                options=list(pageLength=10), 
                rownames=FALSE)})
```

```{r context="render"}
DT::dataTableOutput('DataTable')
```
